---
title: "Steak Preferences in the United States"
subtitle: "Summary Report For the `r params$region` Region"
author: "Author: Melinda K. Higgins, PhD."
date: "March 21, 2018"
output: html_document
params:
  region: 
    label: "Region:"
    value: Mountain
    input: select
    choices: ["East North Central",
    "East South Central", "Middle Atlantic",
    "Mountain", "New England", "Pacific",
    "South Atlantic", "West North Central",
    "West South Central"]
---

```{r setup, include=FALSE}
# define knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

# set any missing NAs in tables to blank
options(knitr.kable.NA='')

# load R packages
library(fivethirtyeight)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)

# load steak_survey dataset
data("steak_survey", package="fivethirtyeight")

# create subset, no missing NAs
# pick a specific region
sdat <- steak_survey %>%
  filter(region==params$region)

```

## Background

In May 2014, Walt Hickey at [538.com](http://fivethirtyeight.com/) published an article entitled ["How Americans Like Their Steak"](https://fivethirtyeight.com/features/how-americans-like-their-steak/). This article utilized survey data collected from 550 people which asked questions related to various risky activities, such as whether they speed in traffic, if they smoke or go skydiving, and if they prefer a riskier lottery. The survey also asked if they eat steak and if so, how they prefer their steak.

## Purpose

The "steak survey" dataset that was used for the ["How Americans Like Their Steak"](https://fivethirtyeight.com/features/how-americans-like-their-steak/) article, is available through the R package [`fivethirtyeight`](https://cran.r-project.org/web/packages/fivethirtyeight/index.html). Using this dataset, this article summarizes the steak preferences for the `r params$region` Region of the United States by gender. 

For our summary presented here, a subset of the original steak survey was extracted to only include responses with no missing data and only included respondents from the `r params$region` region of the US. The data subset used for our summary had `r nrow(sdat)` respondents.

## Demographic Summary Tables - Age

The age categories of the `r nrow(sdat)` survey respondents were:

```{r}
# create table of the age categories
tb <- sdat %>% 
  select(age) %>%
  table() %>%
  prop.table()*100

# use kable from knitr package to make table
# and use kable styling from kableExtra package
knitr::kable(as.data.frame(tb),
             format="html",
             col.names=c("Ages","%"),
             digits=2,
             caption="Ages of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width=FALSE)
```

## Steak Preparation Preference by Gender

Finally, here is a breakdown of the `r nrow(sdat)` survey respondents for the `r params$region` region of the US on how they prefer their steak to be prepared by gender.

```{r}
ggplot(sdat, aes(x = steak_prep, fill = female)) + 
  geom_bar(position="dodge",colour="black") +
  ggtitle(paste0("Steak Preference by Gender: ",
                 params$region," Region")) +
  xlab("Steak Preparation Preference") +
  ylab("Number of Respondents") +
  scale_fill_manual(values=c("skyblue","palevioletred"),
                    name="Gender",
                    breaks=c(FALSE,TRUE),
                    labels=c("Male", "Female")) +
  theme_fivethirtyeight() 
```
