---
title: "Weather Information Preferences in the United States"
subtitle: "Summary Report For the `r params$region` Region"
author: "Author: Melinda K. Higgins, PhD."
date: "`r format(Sys.Date(), '%A, %B %d, %Y')`"
output: html_document
params:
  region: 
    label: "Region:"
    value: Mountain
    input: select
    choices: ["East North Central",
    "East South Central", "Middle Atlantic",
    "Mountain", "New England", "Pacific",
    "South Atlantic", "West North Central",
    "West South Central"]

---

```{r setup, include=FALSE}
# define knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

# set any missing NAs in tables to blank
options(knitr.kable.NA='')

# load R packages
library(fivethirtyeight)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)

# load steak_survey dataset
data("weather_check", package="fivethirtyeight")

# create subset, no missing NAs
# pick a specific region
sdat <- weather_check %>%
  filter(region=="Mountain")

```

## Background

In April 2015, Walt Hickey at [538.com](http://fivethirtyeight.com/) published an article entitled ["Where People Go To Check The Weather"](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/). This article utilized survey data collected from 928 people which asked questions on if they typically check the weather on a daily basis and if so, what source they used to check the weather. Additional follow-up questions were asked about which specific websites or apps they used and if they had a smartwatch if they would check the weather on that smartwatch. 

## Purpose

The "weather check" dataset that was used for the ["Where People Go To Check The Weather"](https://fivethirtyeight.com/features/weather-forecast-news-app-habits/) article, is available through the R package [`fivethirtyeight`](https://cran.r-project.org/web/packages/fivethirtyeight/index.html). Using this dataset, this article summarizes the weather source preferences for the `r params$region` Region of the United States by gender. 
For our summary presented here, a subset of the original weather check dataset was extracted to only include responses with no missing data and only included respondents from the `r params$region` region of the US. The data subset used for our summary had `r nrow(sdat)` respondents.

## Demographics of Survey Respondents in a Merged Table

This section provides demographic summary tables for the `r nrow(sdat)` survey respondents.

```{r}
# create table summary for ages
tb1 <- sdat %>% 
  select(age) %>%
  table() %>%
  prop.table()*100

# create table summary for gender
tb2 <- sdat %>% 
  select(female) %>%
  table() %>%
  prop.table()*100

# create table summary for income
tb3 <- sdat %>% 
  select(hhold_income) %>%
  table() %>%
  prop.table()*100

# convert all tables to data frames
tb1df <- as.data.frame(tb1) 
tb2df <- as.data.frame(tb2)
tb3df <- as.data.frame(tb3)

# merge 1st 2 data frames together
mtb <- merge(data.frame(tb1df, row.names = NULL),
             data.frame(tb2df, row.names = NULL),
             by=0, all=TRUE)[-1]

# merge result with 3rd data frame
mtb2 <- merge(data.frame(mtb, row.names = NULL),
              data.frame(tb3df, row.names = NULL),
              by=0, all=TRUE)[-1]

# use the final data frame
# make into a table with kable
# add styling with kableExtra
# add header with labels spanning 2 columns each
mtb2 %>%
  knitr::kable(format="html",
    col.names=c("Category","%",
                "Category","%",
                "Category","%"),
    digits=2,
    caption="Demographics of Survey Respondents") %>%
  kableExtra::kable_styling(c("striped","bordered"),
                full_width=FALSE) %>%
  add_header_above(c("Ages"=2,"Gender"=2,"Income"=2))
```

## Weather Source Preference by Gender

Finally, here is a breakdown of the `r nrow(sdat)` survey respondents for the `r params$region` region of the US on how they check their daily weather by gender.

```{r}
library(stringr)
ggplot(sdat, 
       aes(x = weather_source, fill = female)) + 
  geom_bar(position="dodge",colour="black") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
  ggtitle(paste0("Weather Source by Gender: ",
                 params$region," Region")) +
  xlab("Weather Source Preference") +
  ylab("Number of Respondents") +
  scale_fill_manual(values=c("skyblue","palevioletred"),
    name="Gender",
    breaks=c(FALSE,TRUE),
    labels=c("Male", "Female")) +
  coord_flip() +
  theme_fivethirtyeight()
```
